Fusion Moves for Correlation Clustering

### 2014-11-07 paper/egpaper_for_review.tex
related work

Segmentation is an important problem in computer vision and data mining as a first step
towards understanding an image or data in general. 
Many algorithms start with an over-segmentation
into superpixels (superregions), which are then clustered into ``perceptually meaningful''
regions.
Usually, the number of these regions is \emph{not} known beforehand.

Recently, the multicut formulation~\cite{chopra_1993_mp, Kappes-2013-multicut}
(sometimes called \emph{correlation clustering}, \cite{bansal_2004_ml})
has become increasingly popular for unsupervised image segmentation.
%
Given an edge-weighted region adjacency graph,
the problem is to find the segmentation which
minimizes the cost of the cut edges.
Such an approach has been shown to yield
state-of-the-art results on the Berkeley Segmentation Database
\cite{andres_2011_iccv,yarkony_2012_eccv,alush_2013_simbad}.

Despite the clear mathematical formulation and nice properties of correlation clustering
to perform a one-shot agglomeration of supervoxels,
the main draw back of correlation clustering is scalability~\cite{nunez_iglesias_2013}
which arise from the NP-hardness of the problem.

@@

Outside computer vision greedy methods has been suggested for correlation clustering problems, see \cite{Elsner-2009} for an overview.

Greedy Algorithms
Step 1: Permute the nodes according a random
Step 2: Assign record to the cluster that minimize the costs. Start a new cluster if costs positive
\begin{itemize}
\item BEST: Cluster containing the most positive edge [Ng et al 2002 ACL]
\item FIRST: Cluster contains the first positive edge [Soon et al 2001 CL]
\item VOTE: Assign to cluster that minimizes objective function. [Elsner et al 08 ACL]
\end{itemize}

PIVOT Algorithm [Ailon et al 2008 JACM]: select pivot element and connect to all positiv neighbord nodes

Run the algorithm for many random permutations , and pick the clustering with
best objective value (better than average run)

BOEM Algorithm [Gioniset al 2007 TKDD]
\begin{itemize}
\item Start with an initial clustering (e.g. output of greedy)
\item Remove one random element from a cluster
\item Make the Best One Element Move (BOEM)
Move it to another cluster or Create a new cluster. 
\end{itemize}

@@

Agglomerative hierarchical clustering (HC) is widely used
in graph / image segmentation
\cite{arbelaez_2006}.

Hierarchical clustering is fast, and 
due to the agglomeration of edge weights, 
hierarchical clustering is robust against noise.
In agglomerative clustering, weights of
parallel edges are merged into single edges.
For image segmentation, the length
weighted mean is used to do this update.

We propose to use Energy Based Hierarchical Clustering (EHC) to find
solutions for correlation clustering.
In each step, the edge with the highest weight $w$ is
contracted (red edges in fig. ~\ref{fig:hc_alg}).
Doing so, parallel edges edges can occur.
To make EHC follow the energy, we propose 
the following update rule:
Whenever there are multiple edges between 
a pair of nodes, these edges are merged into a single edge
by summing up their weights, since we 
minimize the sum of the cut edges.
The algorithm stops if the highest edge weight is
smaller or equal to zero  (green edge in fig. ~\ref{fig:hc_alg}). Any further edge
contraction leads to worse energies.

Given the intrinsic greediness of hierarchical clustering, 
and the fact that EHC does only local decisions, it is 
sensitive against salt and paper noise 
and we do not expect very good solutions.

But EHC is very fast, and can be used to initialize
CGC \cite{beier_2014_cvpr}. Excessive time in CGC
is spend in the \emph{cut phase} to solve the 
first 2 coloring on the complete graph.
As shown in ~\ref{sec:exp},
allowing CGC to start from the EHC solution
can improve performance drastically.

\input{inputs/fig_hc_alg.tex}

>>>>

Correlation clustering~\cite{Bansal-2004} also known as the multicut problem~\cite{chopra_1993_mp} 
is a basic primitive in computer vision~\cite{andres_2011_iccv,alush_2013_simbad} and data mining~\cite{Chierichetti-2014} as a first step
towards understanding an image or data in general. 
The correlation clustering problem is the following:
Given an undirected edge weighted graph, 
where positive weights (\emph{attractive}) encourage the adjacent nodes to be in the same cluster
and negative weights (\emph{repulsive}) encourage the nodes to stay in different cluster, 
find the clustering of nodes such that the sum over weights of edges between clusters is minimized.
Note that this problem formulation does not included a predefined number of clusters.

Despite the clear mathematical formulation and nice properties of correlation clustering
to perform a one-shot agglomeration clustering,
the main draw back of correlation clustering is scalability~\cite{nunez_iglesias_2013}
which arise from the NP-hardness of the problem.

For unsupervised image segmentation algorithms usually start with an over-segmentation
into superpixels (superregions), which are then clustered into ``perceptually meaningful''
regions by correlation clustering on sparse graphs.
Such an approach has been shown to yield
state-of-the-art results on the Berkeley Segmentation Database
\cite{andres_2011_iccv,yarkony_2012_eccv,alush_2013_simbad}.

@@

Outside computer vision greedy methods has been suggested for correlation clustering problems, see \cite{Elsner-2009} for an overview.
A common greedy approach is to randomly permute the nodes and than assign 
iteratively each node to an existing cluster or create a new cluster if costs cannot be decreased by assigning to a cluster.
Common assigning strategies are:
\emph{BEST}; assign to cluster which is linked by the most positive edge [Ng et al 2002 ACL],
\emph{FIRST}; Assign to the cluster which is first linked by a positive edge [Soon et al 2001 CL], and
\emph{VOTE}; Assign to cluster that minimizes objective function. [Elsner et al 08 ACL].

The PIVOT Algorithm [Ailon et al 2008 JACM] iterate over all nodes in random order.
If the node is not assigned it construct a cluster containing the node and all its 
unassigned positively linked neighbors.

Typically these algorithms started for many random permutations, 
and pick the clustering with best objective value.

A widely use post-processing method is  Best One Element Move (BOEM) [Gioniset al 2007 TKDD].
Start with an initial clustering one node is removed from a cluster and reassigned to an existing or new cluster such that the costs are minimized.
BEOM stops if no move can decrease the costs.

@@

Agglomerative hierarchical clustering (HC) is widely used
in graph / image segmentation
\cite{arbelaez_2006}.
Hierarchical clustering is fast, and 
due to the agglomeration of edge weights, 
hierarchical clustering is robust against noise.
In agglomerative clustering, weights of
parallel edges are merged into single edges.
For image segmentation, the length
weighted mean is used to do this update.

We propose to use EHC find
solutions for correlation clustering.
In each step, the edge with the highest weight $w$ is
contracted (red edges in fig. ~\ref{fig:hc_alg}).
Doing so, parallel edges edges can occur.
Whenever there are multiple edges between 
a pair of nodes, these edges are merged into a single edge
by summing up their weights, since we 
minimize the sum of the cut edges.
The algorithm stops if the highest edge weight is
smaller or equal to zero  (green edge in fig. ~\ref{fig:hc_alg}). Any further edge
contraction leads to worse energies.

Given the intrinsic greediness of hierarchical clustering, 
we do not expect very good solutions, but still a good starting 
point for move making algorithm as CGC \cite{beier_2014_cvpr}.
Therefore we propose to initialize  CGC with EBHC
and hope to speedup of CGC with this starting point.

Furthermore, we propose a algorithm to combine / fuse together
multiple EBHC results from perpetuated energies.
This algorithm, namely

\input{inputs/fig_hc_alg.tex} Correlation-Clustering-Fusion (CCF) Moves 
will be discussed in the following section.
